---
title: "HW2_HE"
author: "Hakan Ekiz - IE360 - Fall2020"
date: "10 Ocak 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this project, the Vehicle Price Index dataset is used from EVDS. We will make manipulations and examinations on this dataset. Then, using different datasets as parameters, a model will be developed to estimate the vehicle price index. Data on USD / TL exchange rate, interest rate and the amount of money sent to the treasury from vehicles will be used. Finally, the estimated values and actual values on the model will be drawn on the graph and compared.


## Data

```{r cars, include=FALSE}
library(ggplot2)
library(zoo)
library(data.table)
library(readxl)
library(lubridate)
library(dplyr)
library(forecast)

vehicles = read_xlsx("C:/Users/Z0047JBE/Desktop/dersler/IE360/HW2/fiyatenedeksiarac.xlsx")
vehicles = as.data.table(vehicles[1:156, 1:2])
colnames(vehicles) = c("Date", "Priceindex")
vehicles[,Date := as.yearmon(Date)]
vehicles[,Priceindex := as.numeric(unlist(vehicles$Priceindex))]
vehicles[, Month := month(Date)]
vehicles[, Year := year(Date)-2007]


vehiclesafter2015 = vehicles%>%filter(Year>=8)
vehicles2020 = vehicles%>%filter(Year>=13)


```

The Vehicle Index dataset is taken from EVDS. Date column is reorganized. The dataset is divided to study more clearly. Vehicleafter2015 dataset is created to work on which consists price index values after 2015. 



First, let's look at how this value has changed over time from a general perspective.

```{r}
ggplot(vehiclesafter2015, aes(x = Date, y = Priceindex)) +
  geom_line(size = 1, color="blue") + 
  labs(title = "Priceindex of vehicles in Turkey between 2015-2020 Reported Monthly", 
                             x = "Date",
                             y = "Priceindex") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
ggplot(vehicles2020, aes(x = Month, y = Priceindex)) +
  geom_line(size = 1, color="blue") + 
  labs(title = "Priceindex of vehicles in Turkey in 2020 Reported Monthly", 
                             x = "Month",
                             y = "Priceindex") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```





## Forecasting 

First, we look into the seasonality condition with months. 
```{r}

summary(lm(Priceindex ~ as.factor(Month), vehiclesafter2015))


```


```{r}
checkresiduals(lm(Priceindex ~ as.factor(Month), vehiclesafter2015))

```
Because we have high standart error we should add other factors. 

Now, we add dataset including Income of Turkey Treasury from vehicles. 

```{r}
vehicleincome = read_xlsx("C:/Users/Z0047JBE/Desktop/dersler/IE360/HW2/hazine.xlsx")
vehiclesafter2015[,income :=as.double(unlist(vehicleincome[108:179,2]))]


plot(vehiclesafter2015$income, vehiclesafter2015$Priceindex,  xlab="VehicleIncome", ylab = "Monthly Vehiclepriceindex", main = "Monthly VehiclePriceindex vs monthly vehicle income (2015-2020)")

```

```{r}
summary(lm(Priceindex ~ as.factor(Month) + income, vehiclesafter2015))
```

```{r}
checkresiduals(lm(Priceindex ~ as.factor(Month) + income, vehiclesafter2015))
```
The error rate is lower but still high.


Now, we add dataset including interest rate for purchasing cars. 
```{r}
interestrate = read_xlsx("C:/Users/Z0047JBE/Desktop/dersler/IE360/HW2/vehicleinterestrates.xlsx")
vehiclesafter2015[,interest :=as.double(unlist(interestrate[1:72,2]))]


plot(vehiclesafter2015$interest, vehiclesafter2015$Priceindex,  xlab="Interestrate", ylab = "Monthly Vehiclepriceindex", main = "Monthly VehiclePriceindex vs monthly interestrate (2015-2020)")

```

```{r}
summary(lm(Priceindex ~ as.factor(Month) + income + interest, vehiclesafter2015))
```

```{r}
checkresiduals(lm(Priceindex ~ as.factor(Month) + income + interest, vehiclesafter2015))
```
The error rate is lower


Now, we add dataset including USD/TL error rate. The increase of Dollar generally causes increase in prices. 
```{r}
exchangerate = read_xlsx("C:/Users/Z0047JBE/Desktop/dersler/IE360/HW2/exchangerate.xlsx")
vehiclesafter2015[,exchange :=as.double(unlist(exchangerate[1:72,2]))]


plot(vehiclesafter2015$exchange, vehiclesafter2015$Priceindex,  xlab="ExchangeRate", ylab = "Monthly Vehiclepriceindex", main = "Monthly VehiclePriceindex vs monthly interestrate (2015-2020)")
```

```{r}
summary(lm(Priceindex ~ as.factor(Month) + income + interest + exchange, vehiclesafter2015))

```

```{r}
checkresiduals(lm(Priceindex ~ as.factor(Month) + income + interest + exchange, vehiclesafter2015))

```
The dataset allows to decrease in error rate.

## Model

```{r}
model = lm(Priceindex ~ as.factor(Month) + income + interest + exchange, vehiclesafter2015)
vehiclesafter2015[,prediction:=fitted(model)]
vehiclesafter2015[,residual:=residuals(model)]

plot(vehiclesafter2015$residual, vehiclesafter2015$Priceindex, xlab="Residuals", ylab = "Monthly Prrice Index of vEHİCLES", main = "Monthly Vehicle Price Index vs. Residuals of Proposed Model")

```

Check the model with plotting predicted values and real values. 


```{r}
cols = c("forecast" = "red", "actual" = "blue")
ggplot() +
  geom_line(data=vehiclesafter2015, aes(x=Date, y=prediction, colour="forecast"), lwd=1) +
  geom_line(data=vehiclesafter2015, aes(x=Date, y=Priceindex, colour="actual"), lwd=1) +
  labs(title = "Predicted vs. Actual Number of Monthly Price index of vehicles 2015-2020", 
                             x = "Date",
                             y = "Price Index") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = cols)
```

## References

Dtasets are taken from EVDS. 



